---
title: "relativePosition"
format: 
  html: 
    self-contained: true
    toc: true             
  pdf: 
    number-sections: true 
    toc: true            
    pdf-engine: xelatex 
editor: visual
---

## Quartiles and Interquartile Range

Quartiles divide a sorted data set into four equal parts.

- The lower quartile (Q1) cuts off the lowest 25 percent of values.

- The median (Q2) cuts off the lowest 50 percent of values.

- The upper quartile (Q3) cuts off the lowest 75 percent of values.

The interquartile range (IQR) measures the spread of the middle 50 percent of the data.

$$IQR = Q_3 - Q_1$$

In this example, we will find the interquartile range of eruption duration in the data set faithful.

```{r}
data(faithful)
duration <- faithful$eruptions
summary(duration)
```

### Compute quartiles

```{r}
quantile(duration, c(
  .25, # Q1
  .50, # Median
  .75  # Q3
  ))
```

### Compute interquartile range

```{r}
IQR_duration <- IQR(duration)
IQR_duration
```

The middle 50% of eruption durations spans about `IQR_duration`.

### Box Plot

```{r}

library(ggplot2)

ggplot(data.frame(duration), aes(x = "", y = duration)) +
  geom_boxplot() +
  labs(
    title = "Box Plot of Eruption Duration",
    y = "Duration (minutes)",
    x = NULL
  )
```

## To try on your own

Find the interquartile range of eruption waiting periods. Use the code above as a template.
